---
title: "Project 3"
author: "Jeffrey Zhang and Jaclyn Nguyen"
date: "Monday, February 16, 2015"
output: html_document
---

Project 3 required Jeffrey and I to import a large dataset into Oracle Server and call the data into RStudio for analysis. Below will describe the dataset and the following analysis. 

Dataset: 

Oracle SQL Developer: C##CS329E_JCN565

__1. Loading of R packages: RCurl, ggplot2, dplyr, tidyr (R code not shown)__

```{r, include=FALSE}
library(RCurl)
library (ggplot2)
library(extrafont)
require(dplyr)
require(tidyr)
```

__2. Import dataset from Oracle server__
```{r}
source("../01 data/df_06.R",echo = T)
source("../01 data/df_07.R",echo = T)

```
__3. begin analysis__
First we want to know how many federal funds that each city has. A city could receive the federal funds many times a year and there are also some blank entries (no city names). I did these data wrangling in the following: 

First I plotted the original dataset
```{r}
df_06 %>% select(recipient_city_name,fed_funding_amount_06) %>% ggplot(aes(x=recipient_city_name, y=fed_funding_amount_06,color=recipient_city_name))  + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

I filtered the blank entries and plotted the graph

```{r}
> df_06 %>% select(recipient_city_name,fed_funding_amount_06) %>% filter(recipient_city_name != "") %>% ggplot(aes(x=recipient_city_name, y=fed_funding_amount_06,color=recipient_city_name))  + geom_point() + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

I sum the totol funds for each city for both df_06 and df_07
```{r}
df_06_city_total_fund <- df_06_city_spending %>% select(recipient_city_name) %>% group_by(recipient_city_name) %>% summarize(total_spending = sum(df_06$fed_funding_amount_06)) 
df_07_city_total_fund <- df_07_city_spending %>% select(recipient_city_name) %>% group_by(recipient_city_name) %>% summarize(total_spending = sum(df_07$fed_funding_amount_07)) 
```

Then I inner join two dataset above by city
```{r}
join_by_city <- inner_join(df_06_city_unique_spending,df_07_unique_city_spending,by = "recipient_city_name")
names(join_by_city) <- c("recipient_city_name","total_fed_fund_06","total_fed_fund_07")
```

Use bubble plot to visualize the gragh. First plot bubble plot: fed_fund_06 VS Cities
From the bubble graph, we can easily to see the difference of federal fund amount between these cities by the bubble size
```{r}
join_by_city %>% ggplot(aes(x=recipient_city_name, y=total_fed_fund_06, size = total_fed_fund_06,color=total_fed_fund_06)) + geom_point() + theme(axis.text.x=element_text(angle=80, size=20, vjust=0.5)) + theme(axis.text.x=element_text(size=10, face="bold", vjust=1)) + theme(axis.title.x=element_text(color="forestgreen", vjust=0.35),axis.title.y=element_text(color="cadetblue", vjust=0.35)) + labs(title="Fed_fund_06 VS Cities",y="Total_fund",x="City")
```

Same for 2007
```{r}
join_by_city %>% ggplot(aes(x=recipient_city_name, y=total_fed_fund_07, size = total_fed_fund_07,color=total_fed_fund_07)) + geom_point() + theme(axis.text.x=element_text(angle=80, size=20, vjust=0.5)) + theme(axis.text.x=element_text(size=10, face="bold", vjust=1)) + theme(axis.title.x=element_text(color="forestgreen", vjust=0.35),axis.title.y=element_text(color="cadetblue", vjust=0.35)) + labs(title="Fed_fund_07 VS Cities",y="Total_fund",x="City")
```

continue this part...